---
title: "1_descriptives"
author: "Jackie Kirshenbaum"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document:
    toc: yes
    df_print: paged
  html_notebook:
    toc: yes
    toc_float: yes
editor_options:
  chunk_output_type: console
---

# Libraries

```{r}
library(tidyverse)
library(readxl)
library(scipub)
library(kableExtra)
library(skimr)
library(arsenal)
library(scipub)
library(lubridate)
library(ltm)
library(lmerTest)
library(ggsignif)

theme_set(theme_bw()); theme_update(panel.grid.minor = element_line(linetype = "solid",linewidth = .5),axis.title.x = element_text(face = "bold"),axis.title.y = element_text(face = "bold"))

colorpalette <- c("#F17E0B", "#F5B7B1", "#999999","#BABDE1", "#F5B041", "#B5EAF0", "#CBD1D4",  "#CFE4EF") 

```


# Load data

```{r}
# sleep and si
load("/Volumes/columbia/MAPS_Data/Analysis/Jackie/sleep_and_suicide_2024-2025/weekly_sleep_si_cleaned_October 82025.Rda")
# concurr_sleep_with_si_counts_filt = df with sleep week concurrent to EMA reporting period 
# lead_sleep_with_si_counts_filt = df with future sleep week relative to EMA administration

# sleep and se
load("/../../../../Jackie/sleep_and_suicide_2024-2025/weekly_sleep_se_cleaned_October 82025.Rda")

load("/../../../../../Sleep/sleepdata_April232025.Rda")

# SR
load("/../../../../../../../self_report_07_11_2024.Rda")

# Service use
load("/../../../../../../service_use.Rda")

selfreport <-
  selfreport %>%
  dplyr::mutate(
    study_start_date = SelfReportDate_0
  )
  # dplyr::select(-c(SelfReportDate_0:SelfReportDate_6))

selfreport_itemwise <-
  selfreport_itemwise %>%
  dplyr::select(
    ID, starts_with("psqi"), starts_with("meq"), starts_with("ssi"), starts_with("mfq"), starts_with("shaps"), starts_with("scared")
    )


#### Those with any psych med throughout the study ####
psych_med_ids <-
  su_long %>%
  dplyr::group_by(ID) %>%
  dplyr::mutate(
    any_psych_med = 
      ifelse(
        curr == "1", "1", "0"
    )
  ) %>%
  filter(any_psych_med == "1") %>%
  distinct(ID) # 165 of 223 participants were on any psych med throughout the study (11%)

#### Those with a change from med to non-med (or vice versa) ####
psych_med_change_ids <-
  su_long %>%
  filter(changed == TRUE) %>%
  distinct(ID) # psych med use changed for 35 out of 223 participants (11%)

su <-
  su %>% 
  dplyr::mutate(
    psych_med_change = 
      ifelse(
        ID %in% psych_med_change_ids$ID, 1, 0
      ),
    any_psych_med = 
      ifelse(
        ID %in% psych_med_ids$ID, 1, 0
      )
  ) %>%
  dplyr::select(ID, psych_med_change, any_psych_med)
```

# Merging sleep, SI, and SR

```{r}
## For sleep -> si (lagged df)
merge_si_df_lag <-
  lag_sleep_with_si_counts_filt %>%
  dplyr::select(
    -c(SelfReportDate_0:SelfReportDate_6, study_start_date)
  ) %>%
  left_join(
    selfreport, by = "ID"
  ) %>%
  dplyr::select(
    -c(starts_with("psqi"))
  ) %>%
  left_join(
    su, by = "ID"
  )

merge_si_df_lag <-
  left_join(
    merge_si_df_lag, selfreport_itemwise, by = "ID"
  )
```

# Merging sleep, SE, and SR

```{r}
merge_se_df_lag <-
  left_join(
    weekly_with_stb, selfreport, by = "ID"
  ) %>%
  dplyr::select(
    -c(starts_with("psqi"))
  ) %>%
  left_join(
    su, by = "ID"
  )   

merge_se_df_lag <-
  left_join(
    merge_se_df_lag, selfreport_itemwise, by = "ID"
  )
```

# Selecting variables for SE df

```{r}
merge_se_select <-
  merge_se_df_lag %>%
  dplyr::select(ID, SITE, sex, gender, race, hispanic, age, income, GROUP, SelfReportDate_1:SelfReportDate_6, iPhone, starts_with("psqi"), starts_with("SSI"), sitbi_si_lifetime_i, sitbi_si_pastmonth_i, sitbi_si_pastweek_i, starts_with("MEQ"), starts_with("MFQ"), starts_with("SCARED"), starts_with("SHAPS"), study_start_date, start_date_week_wednesday, schoolclosure, summer, week_num, timeinbed_whole_week_mean:event_cat1_2wk, starts_with("meq"), starts_with("ssi"), starts_with("mfq"), starts_with("shaps"), starts_with("scared"), psych_med_change, any_psych_med) %>%
  dplyr::mutate(
    ID = as.character(ID),
    hispanic = factor(hispanic),
    GROUP = factor(GROUP),
    sex = factor(sex),
    gender = factor(gender),
    iPhone = factor(iPhone),
    schoolclosure = factor(schoolclosure),
    summer = factor(summer),
    psych_med_change = factor(psych_med_change),
    any_psych_med = factor(any_psych_med)
      )

merge_se_select$GROUP <- relevel(merge_se_select$GROUP, ref = "Control")   
merge_se_select$sex <- relevel(merge_se_select$sex, ref = "M")  
merge_se_select$gender <- relevel(merge_se_select$gender, ref = "M") 
```

# Selecting variables for SI df

```{r}
# lagged df
merge_si_select_lag <-
  merge_si_df_lag %>%
  dplyr::select(ID, SITE, sex, gender, race, hispanic, age, income, GROUP, SelfReportDate_0:SelfReportDate_6, iPhone, starts_with("psqi"), starts_with("SSI"), sitbi_si_lifetime_i, sitbi_si_pastmonth_i, sitbi_si_pastweek_i, starts_with("MEQ"), starts_with("MFQ"), starts_with("SCARED"), starts_with("SHAPS"), study_start_date, schoolclosure, summer, week_num, week_start_wednesday:number_of_lead_weeks, timeinbed_whole_week_mean:risk_SA, starts_with("meq"), starts_with("ssi"), starts_with("mfq"), starts_with("shaps"), starts_with("scared"), psych_med_change, any_psych_med) %>%
  dplyr::mutate(
    ID = as.character(ID),
    hispanic = factor(hispanic),
    GROUP = factor(GROUP),
    sex = factor(sex),
    gender = factor(gender),
    iPhone = factor(iPhone),
    psych_med_change = factor(psych_med_change),
    any_psych_med = factor(any_psych_med)    
      )

merge_si_select_lag$GROUP <- relevel(merge_si_select_lag$GROUP, ref = "Control")   
merge_si_select_lag$sex <- relevel(merge_si_select_lag$sex, ref = "M")  
merge_si_select_lag$gender <- relevel(merge_si_select_lag$gender, ref = "M")  

# Observations count
merge_si_select_lag %>%
  dplyr::group_by(ID) %>%
  dplyr::mutate(
    si_3_total = sum(risk_SIsig3_future == 1, na.rm = TRUE),
    si_4_total = sum(risk_SIsig4_future == 1, na.rm = TRUE)
    ) %>%
  dplyr::select(ID, si_3_total, si_4_total) %>%
  distinct() %>%
  dplyr::mutate(
    any_riskSIsig3 = 
      ifelse(
        si_3_total > 0, 1, 0
      ),
    any_riskSIsig4 =
      ifelse(
        si_4_total > 0, 1, 0
      )
  ) %>%
  dplyr::ungroup() %>%
  dplyr::summarise(
    n_sig3 = sum(any_riskSIsig3 == 1),
    n_sig4 = sum(any_riskSIsig4 == 1)
  )

# 66 people had sub-clinical significant SI
# 33 people had clinically significant SI

merge_si_select_lag %>%
  dplyr::group_by(ID) %>%
  dplyr::mutate(
    si_3_total = sum(risk_SIsig3_future == 1, na.rm = TRUE),
    si_4_total = sum(risk_SIsig4_future == 1, na.rm = TRUE)
    ) %>%
  dplyr::select(ID, si_3_total, si_4_total) %>%
  distinct() %>%
  dplyr::ungroup() %>%
  dplyr::summarise(
    n_sig3 = sum(si_3_total),
    n_sig4 = sum(si_4_total)
  )

# 66 people had a total of 224 weekly instances of significant SI
# 33 people had a total of 66 weekly instances of clinically significant SI
33/145 # percentage of those with clinically significant SI
```


## Sample descriptives
Using the larger sample df (`merge_se_df_lag`) for sample descriptives

```{r}
# pulling out person variables
merge_se_select_desc <-
  merge_se_select %>%
  dplyr::ungroup() %>%
  dplyr::select(
    ID, age, sex, gender, race, hispanic, SITE, income, iPhone, GROUP, timeinbed_mean_person, timeinbed_sd_person, bedtime_mean_person, bedtime_sd_person, starts_with("psqi"), starts_with("meq"), starts_with("ssi"), starts_with("mfq"), starts_with("shaps"), starts_with("scared"), psych_med_change, any_psych_med
  ) %>%
  distinct() # 145 pts


# making table 
demo_table_full <-
  FullTable1(
    data = merge_se_select_desc,
    strata = NULL,
    c("age", "sex", "gender", "race", "hispanic", "SITE", "income", "iPhone", "GROUP", "timeinbed_mean_person",  "bedtime_mean_person", "psych_med_change", "any_psych_med"),
    stars = "name", html = F)$table

kable(demo_table_full) %>%
  kable_styling(
    bootstrap_options = c("striped", "hover", "condensed")
    )
write_csv(demo_table_full, "demo_table_n145.csv")

demo_table_by_GROUP <-
  FullTable1(
    data = merge_se_select_desc,
    strata = "GROUP",
    c("age", "sex", "gender", "race", "hispanic", "SITE", "income", "iPhone", "timeinbed_mean_person",  "bedtime_mean_person"),
    stars = "name", html = F)$table

kable(demo_table_by_GROUP) %>%
  kable_styling(
    bootstrap_options = c("striped", "hover", "condensed")
    )
write_csv(demo_table_by_GROUP, "demo_table_n145_byGroup.csv")
```

```{r}
# creating sleep bins based on self report dates for concurrent time-locked data
merge_se_select <- 
  merge_se_select %>%
  dplyr::group_by(ID) %>%
  dplyr::mutate(
    sleep_feature_month = 
      case_when(
        # condition 1
        start_date_week_wednesday >= study_start_date & start_date_week_wednesday < SelfReportDate_1 ~ "bl_to_month_1",
        # condition 2
        start_date_week_wednesday >= study_start_date & is.na(SelfReportDate_1) & is.na(SelfReportDate_3) & is.na(SelfReportDate_6) ~ "bl_onward", # no followups
        # condition 3
        start_date_week_wednesday >= study_start_date & is.na(SelfReportDate_1) & start_date_week_wednesday < SelfReportDate_3 ~ "bl_to_month_3",
        # condition 4
        start_date_week_wednesday >= SelfReportDate_1 & start_date_week_wednesday < SelfReportDate_3 ~ "month_1_to_month_3",
        # condition 5
        start_date_week_wednesday >=SelfReportDate_1 & is.na(SelfReportDate_3) & is.na(SelfReportDate_6) ~ "month_1_onward", # no month 3 / month 6 followup
        # condition 6
        start_date_week_wednesday >= SelfReportDate_1 & is.na(SelfReportDate_3) & start_date_week_wednesday < SelfReportDate_6 ~ "month1_to_month_6",
        # condition 7
        start_date_week_wednesday >= SelfReportDate_3 & start_date_week_wednesday < SelfReportDate_6  ~ "month_3_to_month_6",
        # condition 8
        start_date_week_wednesday >= SelfReportDate_3 & is.na(SelfReportDate_6) ~ "month_3_onward" # no month 6 followup
        ),
    sleep_feature_month = as.factor(sleep_feature_month)
    )  %>%
  dplyr::select(ID:SelfReportDate_6, start_date_week_wednesday, sleep_feature_month, everything()) %>%
  ungroup() 
```

# Internal Consistencies

```{r}
# meq
meq <-
  merge_se_select_desc %>%
  dplyr::select(
    starts_with("meq"), -MEQ_total_i
  ) %>%
  drop_na()
cronbach.alpha(meq, standardized = TRUE, CI = TRUE) # 0.77
# 	missing 2082, 2102 baseline data

# psqi
psqi <-
  merge_se_select_desc %>%
  dplyr::select(
    psqislpqual_c1, psqilaten_c2, psqidurat_c3, psqihse_c4, psqidistb_c5, psqimeds_c6, psqidaydys_c7
  ) %>%
  drop_na()
cronbach.alpha(psqi, standardized = TRUE, CI = TRUE) # 0.672

# ssi
ssi <-
  merge_se_select_desc %>%
  dplyr::select(
    ssi_1_i:ssi_5_i
  ) %>%
  drop_na()
cronbach.alpha(ssi, standardized = TRUE, CI = TRUE) # 0.837

# mfq
mfq <-
  merge_se_select_desc %>%
  dplyr::select(
    mfq_1_i:mfq_13_i
  ) %>%
  drop_na()
cronbach.alpha(mfq, standardized = TRUE, CI = TRUE) # 0.882


# shaps
shaps <-
  merge_se_select_desc %>%
  dplyr::select(
    shaps_1_i:shaps_14_i
  ) %>%
  drop_na()
cronbach.alpha(shaps, standardized = TRUE, CI = TRUE) # 0.88

# scared
scared <-
  merge_se_select_desc %>%
  dplyr::select(
    starts_with("scared")
  ) %>%
  drop_na()
cronbach.alpha(scared, standardized = TRUE, CI = TRUE) # 0.93
```

# Alignment of EARS chronotype: sleep midpoint in relation to self-reported circadian pref 
> in case anyone wants to know

MEQ: the higher the score, the more of a "morning" person you are

## Creating dataframe

```{r}
# sleep features mean bl to 1 mo
ears_sleep_features_bl_to_1mo <-
  merge_se_select %>%
  dplyr::select(ID, start_date_week_wednesday, sleep_feature_month, Midpoint_Roenne_whole_week_mean, Midpoint_Roenne_mean_free_night, timeinbed_whole_week_mean, bedtime_whole_week_mean, risetime_whole_week_mean, MEQ_total_i, starts_with("psqi"), age, sex, iPhone) %>%
  filter(sleep_feature_month == "bl_to_month_1") %>%
  dplyr::group_by(ID) %>%
  dplyr::mutate(
    midpoint_roenne_mean_free_night_bl_to_month_1 = mean(Midpoint_Roenne_mean_free_night),
    midpoint_roenne_mean_bl_to_month_1 = mean(Midpoint_Roenne_whole_week_mean),
    tib_mean_bl_to_month_1 = mean(timeinbed_whole_week_mean),
    bedtime_mean_bl_to_month_1 = mean(bedtime_whole_week_mean),
    risetime_mean_bl_to_month_1 = mean(risetime_whole_week_mean)
  ) %>%
  dplyr::select(ID, sleep_feature_month,  MEQ_total_i, starts_with("psqi"), age, sex, iPhone,  midpoint_roenne_mean_free_night_bl_to_month_1, midpoint_roenne_mean_bl_to_month_1, tib_mean_bl_to_month_1, bedtime_mean_bl_to_month_1, risetime_mean_bl_to_month_1) %>%
  distinct() %>%
  ungroup() %>%
  dplyr::mutate(
    psqidurat = 
      ifelse(
        psqidurat < 150, psqidurat*60, psqidurat
      ),
    psqitib = 
      ifelse(
        psqi_4b_i > 18 & psqi_4b_i < 120, NA_real_, psqi_4b_i
      ),
    psqitib =
      ifelse(
        psqitib < 150, psqitib*60, psqitib
      )
  )

```

## Visualization and Model

```{r}
# sleep midpoint mean bl to one month (CHRONOTYPE / MIDPOINT ON FREE NIGHTS)
ears_sleep_features_bl_to_1mo %>%
  drop_na(midpoint_roenne_mean_free_night_bl_to_month_1) %>% # n=98
  ggplot(
    aes(
      x = MEQ_total_i, y = midpoint_roenne_mean_free_night_bl_to_month_1
    )
  ) +
  geom_point(color = "#F17E0B", alpha = .5) +
  geom_smooth(method = "lm", color = "#F17E0B") +
  labs(
    x = "Self-reported Morningness at BL",
    y = "Midpoint of Sleep on Weekends \nfrom BL to 1-Month"
  )
  
chron_mod <- lm(scale(midpoint_roenne_mean_free_night_bl_to_month_1) ~ scale(MEQ_total_i) + scale(age) + sex + iPhone, data = ears_sleep_features_bl_to_1mo) # sig
summary(chron_mod)
confint(chron_mod, level=0.95)
ggsave("FigureX_chronotype_meq.jpeg", width = 5, height = 4)


# tib mean bl to one month
ears_sleep_features_bl_to_1mo %>%
  ggplot(
    aes(
      x = psqitib, y = tib_mean_bl_to_month_1
    )
  ) +
  geom_point(color = "#F5B7B1", alpha = .5) +
  geom_smooth(method = "lm", color = "#F5B7B1") +
  labs(
    x = "Self-reported TIB at BL",
    y = "TIB \nfrom BL to 1-Month"
  )
  
psqi_mod <- lm(scale(tib_mean_bl_to_month_1) ~ scale(psqitib), data = ears_sleep_features_bl_to_1mo) 
summary(psqi_mod)
confint(psqi_mod, level=0.95)


```

# Histograms for Weekly Sleep Data

## Whole-Week

```{r}
tib_mean <-
  merge_se_select %>%
  ggplot(
    aes(x = timeinbed_whole_week_mean)
  ) +
  geom_histogram() +
  labs(
    x = "Average Weekly Sleep Duration"
  ) +
  theme_classic()

tib_sd <-
  merge_se_select %>%
  ggplot(
    aes(x = timeinbed_whole_week_sd)
  ) +
  geom_histogram() +
  labs(
    x = "Variability of Weekly Sleep Duration"
  ) +
  theme_classic()

bedtime_mean <-
  merge_se_select %>%
  ggplot(
    aes(x = bedtime_whole_week_mean)
  ) +
  geom_histogram() +
  labs(
    x = "Average Weekly Bedtime"
  ) +
  theme_classic()

bedtime_sd <-
  merge_se_select %>%
  ggplot(
    aes(x = bedtime_whole_week_sd)
  ) +
  geom_histogram() +
  labs(
    x = "Variability of Weekly Bedtime"
  )  +
  theme_classic()


descriptives_sleep <- cowplot::plot_grid(tib_mean, tib_sd, bedtime_mean, bedtime_sd, labels = c('A', 'B', 'C', 'D'))
cowplot::save_plot(descriptives_sleep, base_height = 8, base_width = 13, filename = 'Supp_Figure2_descriptives.jpeg')
```

## Weekday / Weekend

```{r}

mean_tib_wkend <- mean(merge_se_select$timeinbed_mean_free_night, na.rm = TRUE)
mean_tib_wkend_label <- as.character(round(mean_tib_wkend, 1))
mean_tib_wkend_label_hrs <- as.character(round(mean_tib_wkend/60, 1))
tib_mean_wkend <-
  merge_se_select %>%
  ggplot(
    aes(x = timeinbed_mean_free_night)
  ) +
  geom_histogram(fill = "#F17E0B") +
  geom_vline(xintercept = mean_tib_wkend) +
  annotate(
    "text", label = paste(mean_tib_wkend_label, "minutes"), x =600, y = 100, size = 4) +
  annotate(
    "text", label = paste(mean_tib_wkend_label_hrs, "hours"), x = 600, y = 90
  ) +
  labs(
    x = "Average Weekend Sleep Duration"
  ) +
  theme_classic()


mean_tib_wkday <- mean(merge_se_select$timeinbed_mean_school_night, na.rm = TRUE)
mean_tib_wkday_label <- as.character(round(mean_tib_wkday, 1))
mean_tib_wkday_label_hrs <- as.character(round(mean_tib_wkday/60, 1))
tib_mean_wkday <-
  merge_se_select %>%
  ggplot(
    aes(x = timeinbed_mean_school_night)
  ) +
  geom_histogram(fill = "#F5B7B1") +
  geom_vline(xintercept = mean_tib_wkday) +
  annotate(
    "text", label = paste(mean_tib_wkday_label, "minutes"), x =600, y = 100, size = 4) +
  annotate(
    "text", label = paste(mean_tib_wkday_label_hrs, "hours"), x = 600, y = 90
  ) +
  labs(
    x = "Average Weekday Sleep Duration"
  ) +
  theme_classic()


sd_tib_wkend <- mean(merge_se_select$timeinbed_sd_free_night, na.rm = TRUE)
sd_tib_wkend_label <- as.character(round(sd_tib_wkend, 1))
sd_tib_wkend_label_hrs <- as.character(round(sd_tib_wkend/60, 1))
tib_sd_wkend <-
  merge_se_select %>%
  ggplot(
    aes(x = timeinbed_sd_free_night)
  ) +
  geom_histogram(fill = "#999999") +
  geom_vline(xintercept = sd_tib_wkend) +
  annotate(
    "text", label = paste(sd_tib_wkend_label, "minutes"), x =200, y = 100, size = 4) +
  annotate(
    "text", label = paste(sd_tib_wkend_label_hrs, "hours"), x = 200, y = 90
  ) +
  labs(
    x = "Variability of Weekend Sleep Duration"
  ) +
  theme_classic()


sd_tib_wkday <- mean(merge_se_select$timeinbed_sd_school_night, na.rm = TRUE)
sd_tib_wkday_label <- as.character(round(sd_tib_wkday, 1))
sd_tib_wkday_label_hrs <- as.character(round(sd_tib_wkday/60, 1))
tib_sd_wkday <-
  merge_se_select %>%
  ggplot(
    aes(x = timeinbed_sd_school_night)
  ) +
  geom_histogram(fill = "#BABDE1") +
  geom_vline(xintercept = sd_tib_wkday) +
  annotate(
    "text", label = paste(sd_tib_wkday_label, "minutes"), x =200, y = 100, size = 4) +
  annotate(
    "text", label = paste(sd_tib_wkday_label_hrs, "hours"), x = 200, y = 90
  ) +
  labs(
    x = "Variability of Weekday Sleep Duration"
  ) +
  theme_classic()


mean_bt_wkend <- mean(merge_se_select$bedtime_mean_free_night, na.rm = TRUE)
mean_bt_wkend_label <- as.character(round(mean_bt_wkend, 1))
mean_bt_wkend_label_hrs <- as.character(round(mean_bt_wkend/60, 1))
bt_mean_wkend <-
  merge_se_select %>%
  ggplot(
    aes(x = bedtime_mean_free_night)
  ) +
  geom_histogram(fill = "#F5B041") +
  geom_vline(xintercept = mean_bt_wkend) +
  annotate(
    "text", label = paste(mean_bt_wkend_label, "minutes"), x = 200, y = 100, size = 4) +
  annotate(
    "text", label = paste(mean_bt_wkend_label_hrs, "hours"), x = 200, y = 90
  ) +
  labs(
    x = "Average Weekend Bedtime"
  ) +
  theme_classic()


mean_bt_wkday <- mean(merge_se_select$bedtime_mean_school_night, na.rm = TRUE)
mean_bt_wkday_label <- as.character(round(mean_bt_wkday, 1))
mean_bt_wkday_label_hrs <- as.character(round(mean_bt_wkday/60, 1))
bt_mean_wkday <-
  merge_se_select %>%
  ggplot(
    aes(x = bedtime_mean_school_night)
  ) +
  geom_histogram(fill = "#CFE4EF") +
  geom_vline(xintercept = mean_bt_wkday) +
  annotate(
    "text", label = paste(mean_bt_wkday_label, "minutes"), x =200, y = 100, size = 4) +
  annotate(
    "text", label = paste(mean_bt_wkday_label_hrs, "hours"), x = 200, y = 90
  ) +
  labs(
    x = "Average Weekday Bedtime"
  ) +
  theme_classic()


sd_bt_wkend <- mean(merge_se_select$bedtime_sd_free_night, na.rm = TRUE)
sd_bt_wkend_label <- as.character(round(sd_bt_wkend, 1))
sd_bt_wkend_label_hrs <- as.character(round(sd_bt_wkend/60, 1))
bt_sd_wkend <-
  merge_se_select %>%
  ggplot(
    aes(x = bedtime_sd_free_night)
  ) +
  geom_histogram(fill = "#B5EAF0") +
  geom_vline(xintercept = sd_bt_wkend) +
  annotate(
    "text", label = paste(sd_bt_wkend_label, "minutes"), x =150, y = 100, size = 4) +
  annotate(
    "text", label = paste(sd_bt_wkend_label_hrs, "hours"), x = 150, y = 90
  ) +
  labs(
    x = "Variability of Weekend Bedtime"
  ) +
  theme_classic()


sd_bt_wkday <- mean(merge_se_select$bedtime_sd_school_night, na.rm = TRUE)
sd_bt_wkday_label <- as.character(round(sd_bt_wkday, 1))
sd_bt_wkday_label_hrs <- as.character(round(sd_bt_wkday/60, 1))
bt_sd_wkday <-
  merge_se_select %>%
  ggplot(
    aes(x = bedtime_sd_school_night)
  ) +
  geom_histogram(fill = "#CBD1D4") +
  geom_vline(xintercept = sd_bt_wkday) +
  annotate(
    "text", label = paste(sd_bt_wkday_label, "minutes"), x =180, y = 100, size = 4) +
  annotate(
    "text", label = paste(sd_bt_wkday_label_hrs, "hours"), x = 180, y = 90
  ) +
  labs(
    x = "Variability of Weekday Bedtime"
  ) +
  theme_classic()


descriptives_sleep_w <- 
  cowplot::plot_grid(tib_mean_wkend, tib_mean_wkday, tib_sd_wkend, tib_sd_wkday, bt_mean_wkend, bt_mean_wkday,  bt_sd_wkend, bt_sd_wkday, labels = c('A', 'B', 'C', 'D', 'E', 'F', 'G', 'H'), ncol = 4)
cowplot::save_plot(descriptives_sleep_w, base_height = 6, base_width = 13, filename = 'Supp_Figure3_descriptives.jpeg')
```

## Weekend / Weekday Differences (weekly level)

```{r}
merge_se_select_w <-
  merge_se_select %>%
  dplyr::select(
    ID, SITE, sex, age, GROUP, event_cat1, MEQ_total_i, start_date_week_wednesday, week_num,  timeinbed_mean_school_night:bedtime_sd_free_night
  ) %>%
  pivot_longer(
    cols = timeinbed_mean_school_night:timeinbed_mean_free_night,
    names_to = "tib_mean",
    values_to = "timeinbed_mean"
  ) %>%
  pivot_longer(
    cols = timeinbed_sd_school_night:timeinbed_sd_free_night,
    names_to = "tib_sd",
    values_to = "timeinbed_sd"
  ) %>%
  dplyr::mutate(
    drop_tib = 
      ifelse(
        (tib_mean == "timeinbed_mean_school_night" & tib_sd == "timeinbed_sd_free_night") |
          (tib_mean == "timeinbed_mean_free_night" & tib_sd == "timeinbed_sd_school_night"),
        "drop",
        NA
      )
  ) %>%
  filter(
    is.na(drop_tib)
  )   %>%
  pivot_longer(
    cols = bedtime_mean_school_night:bedtime_mean_free_night,
    names_to = "bt_mean",
    values_to = "bedtime_mean"
  ) %>%
  pivot_longer(
    cols = bedtime_sd_school_night:bedtime_sd_free_night,
    names_to = "bt_sd",
    values_to = "bedtime_sd"
  ) %>%
  dplyr::mutate(
    drop_bt = 
      ifelse(
        (bt_mean == "bedtime_mean_school_night" & bt_sd == "bedtime_sd_free_night") |
          (bt_mean == "bedtime_mean_free_night" & bt_sd == "bedtime_sd_school_night"),
        "drop",
        NA
      )
  ) %>%
  filter(
    is.na(drop_bt)
  )  %>%
   dplyr::mutate(
    drop = 
      ifelse(
        (tib_mean == "timeinbed_mean_school_night" & bt_mean == "bedtime_mean_free_night") |
          (tib_mean == "timeinbed_mean_free_night" & bt_mean == "bedtime_mean_school_night"),
        "drop",
        NA
      )
  ) %>%
  filter(
    is.na(drop)
  )  %>%
  dplyr::select(-c(drop_tib, drop_bt, drop)) %>%
  dplyr::mutate(
    wknight =
      ifelse(
       str_detect(tib_mean, "school_night"), "school", "weekend"
      )
  ) %>%
  dplyr::select(-c(tib_mean, tib_sd, bt_mean, bt_sd)) %>%
  dplyr::mutate(
    GROUP_re = 
      ifelse(
        GROUP == "Ideator" | GROUP == "Attempter", "STB", "CTL"
      )
  )

summary(lmer(timeinbed_mean ~ wknight + (1 | ID), data = merge_se_select_w))
# no difference in average time in bed on weekends and weekdays

summary(lmer(timeinbed_sd ~ wknight  + (1 | ID), data = merge_se_select_w))
# there is more variability of time in bed on weeknights (school nights) compared to weekends

summary(lmer(bedtime_mean ~ wknight + (1 | ID), data = merge_se_select_w))
# youth go to bed later on the weekends compared to weeknights

summary(lmer(bedtime_sd ~ wknight + (1 | ID), data = merge_se_select_w))
# there is more variability of bedtime on weeknights (school nights) compared to weekends

# average tib
avg_tib_diffs <-
  merge_se_select_w %>%
  # slice_sample(n=800, replace = TRUE) %>%  
  ggplot(
    aes(
      x = week_num, y = timeinbed_mean, group = ID, color = wknight
    )
  ) +
  geom_line(alpha = .5) +
  geom_smooth(
    method = "lm",
    aes(x = week_num, y = timeinbed_mean, group = wknight, color = wknight)
  ) +
  scale_color_manual(values = c("#F5B041", "#BABDE1")) +
  theme_classic() +
  labs(
    x = "Week",
    y = "Average Sleep Duration",
    color = "Weeknight"
  )
  
# tib sd
sd_tib_diffs <-
  merge_se_select_w %>%
  # slice_sample(n=800, replace = TRUE) %>%  
  ggplot(
    aes(
      x = week_num, y = timeinbed_sd, group = ID, color = wknight
    )
  ) +
  geom_line(alpha = .5) +
  geom_smooth(
    method = "lm",
    aes(x = week_num, y = timeinbed_sd, group = wknight, color = wknight)
  ) +
  scale_color_manual(values = c("#F5B041", "#BABDE1")) +
  theme_classic() +
  labs(
    x = "Week",
    y = "Variability of Sleep Duration",
    color = "Weeknight"
  )

# bedtime mean
avg_bt_diffs <-
  merge_se_select_w %>%
  # slice_sample(n=800, replace = TRUE) %>%  
  ggplot(
    aes(
      x = week_num, y = bedtime_mean, group = ID, color = wknight
    )
  ) +
  geom_line(alpha = .5) +
  geom_smooth(
    method = "lm",
    aes(x = week_num, y = bedtime_mean, group = wknight, color = wknight)
  ) +
  scale_color_manual(values = c("#F5B041", "#BABDE1")) +
  theme_classic() +
  labs(
    x = "Week",
    y = "Average Bedtime",
    color = "Weeknight"
  )

# bedtime sd
sd_bt_diffs <-
  merge_se_select_w %>%
  # slice_sample(n=800, replace = TRUE) %>%  
  ggplot(
    aes(
      x = week_num, y = bedtime_sd, group = ID, color = wknight
    )
  ) +
  geom_line(alpha=.5) +
  geom_smooth(
    method = "lm",
    aes(x = week_num, y = bedtime_sd, group = wknight, color = wknight)
  ) +
  scale_color_manual(values = c("#F5B041", "#BABDE1")) +
  theme_classic() +
  labs(
    x = "Week",
    y = "Variability of Bedtime",
    color = "Weeknight"
  )
descriptives_sleep_diffs <- 
  cowplot::plot_grid(avg_tib_diffs, sd_tib_diffs, avg_bt_diffs, sd_bt_diffs, labels = c('A', 'B', 'C', 'D'), ncol = 2)
cowplot::save_plot(descriptives_sleep_diffs, base_height = 6, base_width = 10, filename = 'Supp_Figure4_descriptives.jpeg')
```

## Weekend / Weekday Differences (daily level)

```{r}
daily_sleep_single <-
  daily_sleep %>%
  dplyr::group_by(ID) %>% 
  dplyr::mutate(
    dayX = wday(dt_feature, week_start = getOption("lubridate.week.start", 3))
  ) %>%
  dplyr::select(ID, dt_feature, wday_name, dayX, everything()) %>%
  dplyr::mutate(
    dt_feature = as.Date(dt_feature),
    start_date = 
      floor_date(dt_feature, unit = "week", week_start = getOption("lubridate.week.start", 3)),
    end_date = start_date + 7
  ) %>%
  dplyr::select(ID, dt_feature, wday_name, dayX, start_date, everything()) %>%
  distinct() %>%
  dplyr::group_by(ID) %>%
  dplyr::mutate(
    week_num_sleep_lead = as.numeric(dt_feature - min(start_date))%/%7,
    week_num_sleep_lag = as.numeric(dt_feature - min(end_date))%/%7
  ) %>%
  dplyr::select(ID, dt_feature, wday_name, dayX, start_date, end_date, week_num_sleep_lead, week_num_sleep_lag, everything()) %>%
  arrange(ID, dt_feature) %>%
  dplyr::mutate(
    n_total_daily_motion_obs = n()
  ) %>%
  ungroup() %>%
  group_by(ID, week_num_sleep_lead) %>%
  dplyr::mutate(
    n_total_motion_this_week = n()
  ) %>%
  ungroup() %>%
  group_by(ID) %>%
  dplyr::mutate(
    week_num_sleep_lag = factor(week_num_sleep_lag),
    week_num_sleep_lead = factor(week_num_sleep_lead),
    num_wks_with_motion = length(unique(week_num_sleep_lead))
  ) %>%
  ungroup()  # 150 pts with 15600 obs
    

weekly_sleep_filtered_sum <-
  daily_sleep_single %>%
  filter(
    n_total_motion_this_week > 1
  ) %>%
  group_by(ID) %>%
  dplyr::summarise(
    num_wks_with_more_than_1_obs_motion = length(unique(week_num_sleep_lead))
  )

daily_sleep_single <-
  left_join(daily_sleep_single, weekly_sleep_filtered_sum, by = c("ID")) %>%
  filter(
    n_total_motion_this_week > 1
    )  %>%
  dplyr::mutate(
    wday_name = lubridate::wday(dt_feature, label = TRUE),
    night_type = 
      ifelse(
        wday_name == "Fri" | wday_name == "Sat", "free_night", "school_night"
        )
    )
```


```{r}
# Save
save(file = format(Sys.Date(), 'weekly_sleep_and_stb_and_sr_cleaned_%B%e%Y.Rda'), list=c("merge_se_select", "merge_se_select_desc", "merge_si_select_lag"))
```


# References

::: {#refs custom-style="Bibliography"}
:::
