---
title: "2_sleep_and_se_associations"
author: "Jackie Kirshenbaum"
date: "`r format(Sys.Date(), '%B %e %Y')`"
output:
  html_document:
    df_print: paged
  toc: yes
  toc_float: yes
---

# Libraries and Themes

```{r}
library(tidyverse)
library(scipub)
library(kableExtra)
library(skimr)
library(lmerTest)
library(brms)
library(naniar)
library(sjPlot)
library(broom)
library(ggeffects)
library(patchwork)
library(jtools)
library(interactions)
library(broom)
library(broom.mixed)
library(tidymodels)
library(dotwhisker)


colorpalette <- c("#F17E0B", "#F5B7B1", "#999999","#BABDE1", "#F5B041", "#B5EAF0", "#CBD1D4",  "#CFE4EF") 
```

# Load data

```{r}
# weekly with suicide event data
load("/../../../../Jackie/sleep_and_suicide_2024-2025/weekly_sleep_and_stb_and_sr_cleaned_October 82025.Rda")

load("/../../../../../../../self_report_07_11_2024.Rda")

load("/../../../../../../service_use.Rda")

```

# Did psychiatric medication change?

```{r}
#### Those with any psych med throughout the study ####
psych_med_ids <-
  su_long %>%
  dplyr::group_by(ID) %>%
  dplyr::mutate(
    any_psych_med = 
      ifelse(
        curr == "1", "1", "0"
    )
  ) %>%
  filter(any_psych_med == "1") %>%
  distinct(ID) # 165 of 223 participants were on any psych med throughout the study (11%)

#### Those with a change from med to non-med (or vice versa) ####
psych_med_change_ids <-
  su_long %>%
  filter(changed == TRUE) %>%
  distinct(ID) # psych med use changed for 35 out of 223 participants (11%)

#### joining psych med use with main df (those with sleep data) ####
merge_se_select <- 
  merge(
    merge_se_select, su, by = "ID"
  ) %>%
  dplyr::mutate(
    psych_med_change = 
      ifelse(
        ID %in% psych_med_change_ids$ID, 1, 0
      ),
    any_psych_med = 
      ifelse(
        ID %in% psych_med_ids$ID, 1, 0
      )
  )

psych_med_ids <-
  merge_se_select %>%
  filter(any_psych_med == 1) %>%
  distinct(ID) # # For those who have sleep data, 115 our of 145 participants had any psych med use (80%)

psych_med_change_ids <-
  merge_se_select %>%
  filter(psych_med_change == 1) %>%
  distinct(ID) # For those who have sleep data, psych med use changed for 24 out of 145 participants (17%)
```

# Pulling out missigness variable from SI df

```{r}
weekly_si_missingness <-
  merge_si_select_lag %>%
  dplyr::select(ID, missing_lead_weeks) %>%
  distinct()

# joining with se df
merge_se_select <- merge(merge_se_select, weekly_si_missingness, by = "ID")

```

# Scaling variables

```{r}
merge_se_select_scaled <- 
  merge_se_select %>%
  dplyr::ungroup() %>%
  dplyr::group_by(ID) %>% # p-mean centering (creating within-person / idiographic variables)
  dplyr::mutate(
    # tib weekly
    timeinbed_whole_week_mean_pmean = as.numeric(scale(timeinbed_whole_week_mean)), # person's mean duration that week relative to their mean duration on other weeks 
    timeinbed_whole_week_mean_pmean_hrs = as.numeric(scale(timeinbed_whole_week_mean/60, scale = FALSE)),
    timeinbed_mean_school_night_pmean = as.numeric(scale(timeinbed_mean_school_night)),
    timeinbed_mean_school_night_pmean_hrs = as.numeric(scale(timeinbed_mean_school_night/60, scale = FALSE)),    
    timeinbed_mean_free_night_pmean = as.numeric(scale(timeinbed_mean_free_night)),
    timeinbed_mean_free_night_pmean_hrs = as.numeric(scale(timeinbed_mean_free_night/60, scale = FALSE)),    
    timeinbed_whole_week_sd_pmean = as.numeric(scale(timeinbed_whole_week_sd)),   # person's variability in duration that week relative to their mean variability in duration on other weeks
    timeinbed_whole_week_sd_pmean_hrs = as.numeric(scale(timeinbed_whole_week_sd/60, scale = FALSE)), 
    timeinbed_sd_school_night_pmean = as.numeric(scale(timeinbed_sd_school_night)),
    timeinbed_sd_school_night_pmean_hrs = as.numeric(scale(timeinbed_sd_school_night/60, scale = FALSE)),    
    timeinbed_sd_free_night_pmean = as.numeric(scale(timeinbed_sd_free_night)),
    timeinbed_sd_free_night_pmean_hrs = as.numeric(scale(timeinbed_sd_free_night/60, scale = FALSE)),    
    # bedtime weekly
    bedtime_whole_week_mean_pmean = as.numeric(scale(bedtime_whole_week_mean)),  
    bedtime_mean_school_night_pmean = as.numeric(scale(bedtime_mean_school_night)),
    bedtime_mean_free_night_pmean = as.numeric(scale(bedtime_mean_free_night)),
    bedtime_whole_week_sd_pmean = as.numeric(scale(bedtime_whole_week_sd)),
    bedtime_sd_school_night_pmean = as.numeric(scale(bedtime_sd_school_night)),
    bedtime_sd_free_night_pmean = as.numeric(scale(bedtime_sd_free_night)),
    # mood weekly
    mood_whole_week_mean_pmean = as.numeric(scale(mood_whole_week_mean)),  
    mood_mean_school_night_pmean = as.numeric(scale(mood_mean_school_night)),
    mood_mean_free_night_pmean = as.numeric(scale(mood_mean_free_night))
  ) %>%
  dplyr::ungroup() %>% 
  dplyr::mutate(
    # sex
    sex = factor(sex),
    # age
    age_z = as.numeric(scale(age)),
    week_num_time_z_rescaled = as.numeric(scale(week_num)),
    schoolclosure = factor(schoolclosure),
    summer = factor(summer),
    psych_med_change = factor(psych_med_change),
    any_psych_med = factor(any_psych_med),
    # SR variables
    SSI5_i_z = as.numeric(scale(SSI5_i)),
    MFQ_i_z = as.numeric(scale(MFQ_i)),
    SCARED_i_z = as.numeric(scale(SCARED_i)),
    SHAPS_i_z = as.numeric(scale(SHAPS_TOT_r_i)),
    MEQ_total_i_z = as.numeric(scale(MEQ_total_i)),
    psqitotal_z = as.numeric(scale(psqitotal)),
    weekly_missingness_z = as.numeric(scale(missing_lead_weeks))
  ) %>%
  dplyr::mutate(across(ends_with("z"), ~ifelse(is.nan(.), NA, .)))
```

# Covariates?

```{r}
cov_age_sex <- glmer(event_cat1 ~ age_z + sex + (1 | ID), family=binomial,  na.action=na.omit, data = merge_se_select_scaled)
cov_sch_close <- glmer(event_cat1 ~ schoolclosure + (1 | ID), family=binomial,  na.action=na.omit, data = merge_se_select_scaled) 
cov_summ <- glmer(event_cat1 ~ summer + (1 | ID), family=binomial,  na.action=na.omit, data = merge_se_select_scaled)
cov_week <- glmer(event_cat1 ~ week_num_time_z_rescaled + (1 | ID), family=binomial,  na.action=na.omit, data = merge_se_select_scaled) 
cov_phone <- glmer(event_cat1 ~ iPhone +  (1 | ID), family=binomial,  na.action=na.omit, data = merge_se_select_scaled)
cov_ssi <- glmer(event_cat1 ~  SSI5_i_z + (1 | ID), family=binomial,  na.action=na.omit, data = merge_se_select_scaled)
cov_mfq <- glmer(event_cat1 ~ MFQ_i_z +  (1 | ID), family=binomial,  na.action=na.omit, data = merge_se_select_scaled)
cov_meq <- glmer(event_cat1 ~ MEQ_total_i_z +   (1 | ID), family=binomial,  na.action=na.omit, data = merge_se_select_scaled)
cov_scared <- glmer(event_cat1 ~ SCARED_i_z + (1 | ID), family=binomial,  na.action=na.omit, data = merge_se_select_scaled)
cov_shaps <- glmer(event_cat1 ~ SHAPS_i_z + (1 | ID), family=binomial,  na.action=na.omit, data = merge_se_select_scaled)
cov_psqi <- glmer(event_cat1 ~ psqitotal_z + (1 | ID), family=binomial,  na.action=na.omit, data = merge_se_select_scaled)
cov_mood <- glmer(event_cat1 ~ mood_whole_week_mean_pmean + (1 | ID), family=binomial,  na.action=na.omit, data = merge_se_select_scaled)
cov_mood_school <- glmer(event_cat1 ~ mood_mean_school_night_pmean + (1 | ID), family=binomial,  na.action=na.omit, data = merge_se_select_scaled)
cov_mood_free <- glmer(event_cat1 ~ mood_mean_free_night_pmean + (1 | ID), family=binomial,  na.action=na.omit, data = merge_se_select_scaled)

tab_model(cov_age_sex, cov_sch_close, cov_summ, cov_week, cov_phone, cov_ssi, cov_mfq, cov_meq, cov_scared, cov_shaps, cov_psqi, cov_mood, cov_mood_school, cov_mood_free)
# week number is the only covariate associated with SE

# Looking at psych meds within those who had SE
se_only <- 
  merge_se_select_scaled %>% 
  dplyr::select(ID, SelfReportDate_1:start_date_week_wednesday, study_start_date, sleep_feature_month, psych_med_change, any_psych_med, event_cat1) %>%
  filter(event_cat1 == 1) %>% 
  dplyr::select(
    ID, psych_med_change:event_cat1
  ) %>%
  distinct(ID, .keep_all = TRUE)
```


##### Supplementary CR Analysis of Missingness #####

```{r}
merge_se_select_scaled_cr <-
  merge_se_select_scaled %>%
  dplyr::group_by(ID) %>%
  dplyr::mutate(
    anySE = 
      ifelse(event_cat1 == 1, "1", "0"),
    anySE = as.factor(anySE)
  ) %>%
  dplyr::select(
    ID, weekly_missingness_z, missing_lead_weeks, anySE, SSI5_i_z, SHAPS_i_z, MFQ_i_z
  ) %>%
  distinct()

t.test(missing_lead_weeks ~ anySE, data = merge_se_select_scaled_cr)

missingness_m1 <- glmer(event_cat1 ~ weekly_missingness_z + (1 | ID), family=binomial,  na.action=na.omit, data = merge_se_select_scaled)
summary(missingness_m1)
tab_model(missingness_m1, show.std = TRUE)

missingness_m2 <- lm(weekly_missingness_z  ~ SSI5_i_z, na.action=na.omit, data = merge_se_select_scaled_cr)
summary(missingness_m2)
tab_model(missingness_m2, show.std = TRUE)

missingness_m3 <- lm(weekly_missingness_z  ~ SHAPS_i_z, na.action=na.omit, data = merge_se_select_scaled_cr)
summary(missingness_m3)
tab_model(missingness_m3, show.std = TRUE)

missingness_m4 <- lm(weekly_missingness_z  ~ MFQ_i_z, na.action=na.omit, data = merge_se_select_scaled_cr)
summary(missingness_m4)
tab_model(missingness_m4, show.std = TRUE)
```


# Person-Centered Logistic Reg Whole Week Mean + SD

```{r}
#### TIB (whole week)
log_tib_step1 <- glmer(event_cat1 ~ timeinbed_whole_week_sd_pmean + timeinbed_whole_week_mean_pmean  +  week_num_time_z_rescaled + (1 | ID), family=binomial(link = 'logit'), data = merge_se_select_scaled)
summary(log_tib_step1) # sig main effect of TIB

tab_model(log_tib_step1, show.std = TRUE, pred.labels = c("Intercept", "Weekly TIB SD (pmean)", "Weekly TIB Mean (pmean)", "Weeks Since Baseline"))


#### Supplementary analysis looking at a 2-week lag sleep -> SE ####
log_tib_step1_supp <- glmer(event_cat1_2wk ~ timeinbed_whole_week_sd_pmean + timeinbed_whole_week_mean_pmean  +  week_num_time_z_rescaled + (1 | ID), family=binomial(link = 'logit'), data = merge_se_select_scaled)
summary(log_tib_step1_supp)
tab_model(log_tib_step1_supp, show.std = TRUE, pred.labels = c("Intercept", "Weekly TIB SD (pmean)", "Weekly TIB Mean (pmean)", "Weeks Since Baseline"))


#### BEDTIME (whole week)
log_on_step1 <- glmer(event_cat1 ~ bedtime_whole_week_sd_pmean + bedtime_whole_week_mean_pmean +  week_num_time_z_rescaled + (1 | ID), family=binomial(link = 'logit'),  data = merge_se_select_scaled)
summary(log_on_step1) 

tab_model(log_on_step1, show.std = TRUE, pred.labels = c("Intercept", "Weekly Bedtime SD (pmean)", "Weekly Bedtime Mean (pmean)", "Weeks Since Baseline"))
```


###################### WEEKEND / WEEKDAY ############################

## TIB Weekdays

```{r}
log_tib_step1_ww <- glmer(event_cat1 ~ timeinbed_mean_school_night_pmean + timeinbed_sd_school_night_pmean + week_num_time_z_rescaled + (1 | ID), family=binomial(link = 'logit'), data = merge_se_select_scaled)
summary(log_tib_step1_ww) # sig main effect of TIB

tab_model(log_tib_step1_ww, show.std = TRUE, pred.labels = c("Intercept", "Weekday Sleep Duration Mean (pmean)", "Weekday Sleep Duration SD (pmean)", "Weeks Since Baseline"))

#### Centered/Unscaled ####
log_tib_step1_ww_unsc <- glmer(event_cat1 ~ timeinbed_mean_school_night_pmean_hrs + timeinbed_sd_school_night_pmean_hrs + week_num_time_z_rescaled + (1 | ID), family=binomial(link = 'logit'), data = merge_se_select_scaled)
summary(log_tib_step1_ww_unsc) 

tab_model(log_tib_step1_ww_unsc, show.std = TRUE, pred.labels = c("Intercept", "Weekday Sleep Duration Mean (pmean)", "Weekday Sleep Duration SD (pmean)", "Weeks Since Baseline"))
# For every hour increase in average sleep duration on weekdays, there is a near two-fold increase in the likelihood of a suicide event the next week
```

#### TIB plot

```{r}
plot_model(log_tib_step1, type = "est", axis.title = "Odds Ratio", show.values = TRUE, show.p = TRUE) +
  scale_y_log10(limits = c(0.25, 5))
plot_model(log_tib_step1_ww, type = "est", axis.title = "Odds Ratio", show.values = TRUE, show.p = TRUE, vline.color = 'black', colors = "hero") +
  scale_y_log10(limits = c(0.25, 8)) +
  theme_classic()
ggsave("FigureX_tib_weekday_mean_and_SE_OR.jpeg", width = 4.5, height = 3)
```

## TIB Weekends

```{r}
log_tib_step1_wwend <- glmer(event_cat1 ~ timeinbed_mean_free_night_pmean + timeinbed_sd_free_night_pmean + week_num_time_z_rescaled + (1 | ID), family=binomial(link = 'logit'), data = merge_se_select_scaled)
summary(log_tib_step1_wwend) 

tab_model(log_tib_step1_wwend,  show.std = TRUE, pred.labels = c("Intercept", "Weekend Sleep Duration Mean (pmean)", "Weekend Sleep Duration SD (pmean)", "Weeks Since Baseline"))
```

## Bedtime Weekdays

```{r}
log_on_step1_ww <- glmer(event_cat1 ~ bedtime_mean_school_night_pmean + bedtime_sd_school_night_pmean + week_num_time_z_rescaled + (1 | ID), family=binomial(link = 'logit'), data = merge_se_select_scaled)
summary(log_on_step1_ww) 

tab_model(log_on_step1_ww,  show.std = TRUE, pred.labels = c("Intercept", "Weekday Bedtime Mean (pmean)", "Weekday Bedtime SD (pmean)", "Weeks Since Baseline"))
```

## Bedtime Weekends

```{r}
log_on_step1_wwend <- glmer(event_cat1 ~ bedtime_mean_free_night_pmean + bedtime_sd_free_night_pmean + week_num_time_z_rescaled + (1 | ID), family=binomial(link = 'logit'), data = merge_se_select_scaled)
summary(log_on_step1_wwend) 

tab_model(log_on_step1_wwend, show.std = TRUE, pred.labels = c("Intercept", "Weekend Bedtime Mean (pmean)", "Weekend Bedtime SD (pmean)", "Weeks Since Baseline"))
```

# Overall Models

```{r}
tab_model(log_tib_step1, log_tib_step1_ww, log_tib_step1_wwend, show.std = TRUE, pred.labels = c("Intercept", "Whole-Week Sleep Duration SD (pmean)", "Whole-Week Sleep Duration Mean (pmean)", "Weeks Since Baseline", "Weekday Sleep Duration Mean (pmean)", "Weekday Sleep Duration SD (pmean)", "Weekend Sleep Duration Mean (pmean)", "Weekend Sleep Duration SD (pmean)"))

tab_model(log_on_step1, log_on_step1_ww, log_on_step1_wwend, show.std = TRUE, pred.labels = c("Intercept", "Whole-Week Bedtime SD (pmean)", "Whole-Week Bedtime Mean (pmean)", "Weeks Since Baseline", "Weekday Bedtime Mean (pmean)", "Weekday Bedtime SD (pmean)", "Weekend Bedtime Mean (pmean)", "Weekend Bedtime SD (pmean)"))
```

Overall, results indicate that a greater time-in-bed relative to one's typical time-in-bed on a given a week is related to next week presence of a SE. This seems to be driven by time-in-bed during weekdays.

#### OR Plot
```{r}
# Option 1

m1_or_se <- broom::tidy(log_tib_step1, conf.int=TRUE, conf.level = 0.95, exponentiate=TRUE)  %>%
  filter(term %in% c("timeinbed_whole_week_mean_pmean", "timeinbed_whole_week_sd_pmean")) %>%
  dplyr::select(term, estimate, std.error, p.value, conf.low, conf.high) %>%
  dplyr::mutate(., sig = ifelse((conf.low > 1 & conf.high > 1) |
                                            (conf.low < 1 & conf.high < 1), '*', ''))

m2_or_se <- broom::tidy(log_tib_step1_ww, conf.int=TRUE, conf.level = 0.95, exponentiate=TRUE)  %>%
  filter(term %in% c("timeinbed_mean_school_night_pmean", "timeinbed_sd_school_night_pmean")) %>%
  dplyr::select(term, estimate, std.error, p.value, conf.low, conf.high) %>%
  dplyr::mutate(., sig = ifelse((conf.low > 1 & conf.high > 1) |
                                            (conf.low < 1 & conf.high < 1), '*', ''))

m3_or_se <- broom::tidy(log_tib_step1_wwend, conf.int=TRUE, conf.level = 0.95, exponentiate=TRUE)  %>%
  filter(term %in% c("timeinbed_mean_free_night_pmean", "timeinbed_sd_free_night_pmean")) %>%
  dplyr::select(term, estimate, std.error, p.value, conf.low, conf.high) %>%
  dplyr::mutate(., sig = ifelse((conf.low > 1 & conf.high > 1) |
                                            (conf.low < 1 & conf.high < 1), '*', ''))

comb_models <- rbind(m1_or_se, m2_or_se, m3_or_se)

or_se_plot <- 
  comb_models %>%
    dplyr::mutate(
      term = dplyr::recode(
        term, 
        'timeinbed_whole_week_mean_pmean' = 'Within-Person Average \n Sleep Duration \n during the Whole Week',
        'timeinbed_whole_week_sd_pmean' = 'Within-Person Variability \n of Sleep Duration \n during the Whole Week',    
        'timeinbed_mean_school_night_pmean' = 'Within-Person Average \n Sleep Duration \n  on Weekdays',
        'timeinbed_sd_school_night_pmean' = 'Within-Person Variability \n of Sleep Duration \n on Weekdays',
        'timeinbed_mean_free_night_pmean' = 'Within-Person Average \n Sleep Duration \n  on Weekends',
        'timeinbed_sd_free_night_pmean' = 'Within-Person Variability \n of Sleep Duration \n on Weekends'
        )
      ) %>%
    ggplot(
      data=., 
      aes(x = term, y = estimate)
      ) +
  geom_errorbar(
    aes(ymin = conf.low, ymax = conf.high),
    lwd = .3, width = 0, position = position_dodge(0.6)
    ) +
  geom_point(position = position_dodge(0.6), aes(color = term), size = 3, show.legend = FALSE) +
  labs(x = NULL, y = "Adjusted Odds Ratio", title = 'Within-Person Suicidal Event Models') +
  geom_hline(yintercept = 1, lty = 2) +
  geom_text(aes(label = sig, y = conf.high*1.02), size = 5) +
  labs(color = NULL) +
  scale_color_manual(values = colorpalette) +
  theme(
    strip.text.y = element_blank(),
    axis.text.x = element_text(colour = "black"),
    axis.text.y = element_text(colour = "black"),
    legend.position = "none"
    )  +
  scale_y_log10() +
  scale_y_continuous(limits = c(0.30, 5), breaks=seq(0.30,5,.70)) +
  coord_flip() +
  theme_classic()
or_se_plot
ggsave("or_se_plot.jpeg", width = 5, height = 4)
```

## Sensitivity Analyses

### sensitivity: random removal of events for pt who had 6 events

```{r}
set.seed(123)
sum_subjs <-
  merge_se_select_scaled %>%
  dplyr::select(
    ID, age, sex, GROUP, week_num, event_cat1
  ) %>%
  dplyr::group_by(ID) %>%
  dplyr::mutate(
    sum_events = sum(event_cat1)
  ) # 1082 random removal of 4 events


merge_se_select_scaled_1082 <-
  merge_se_select_scaled %>%
  filter(ID == "1082" & event_cat1 == 1) %>%
  slice_sample(n=2)

merge_se_select_scaledx <-
  merge_se_select_scaled %>%
  filter(
    ID != "1082"
  )

merge_se_select_scaledxx <- 
  rbind(
    merge_se_select_scaledx, merge_se_select_scaled_1082
  )

# random sample method
log_tib_step1_ww_sens <- glmer(event_cat1 ~ timeinbed_mean_school_night_pmean + timeinbed_sd_school_night_pmean + week_num_time_z_rescaled + (1 | ID), family=binomial(link = 'logit'), data = merge_se_select_scaledxx)
summary(log_tib_step1_ww_sens) 

tab_model(log_tib_step1_ww_sens,  show.std = TRUE, pred.labels = c("Intercept", "Weekday Sleep Duration Mean (pmean)", "Weekday Sleep Duration SD (pmean)", "Weeks Since Baseline"))
```

### univariate analyses

```{r}
#### TIB (whole week)
log_tib_step1_uni <- glmer(event_cat1 ~  timeinbed_whole_week_mean_pmean + (1 | ID), family=binomial(link = 'logit'), data = merge_se_select_scaled)
summary(log_tib_step1_uni) # sig main effect of TIB
tab_model(log_tib_step1_uni,  show.std = TRUE, pred.labels = c("Intercept", "Weekly Sleep Duration Mean (pmean)")) # OR = 1.80, CI 1.03-3.07, p=.039

log_tib_step1_ww_uni <- glmer(event_cat1 ~ timeinbed_mean_school_night_pmean + (1 | ID), family=binomial(link = 'logit'), data = merge_se_select_scaled)
summary(log_tib_step1_ww_uni) 
tab_model(log_tib_step1_ww_uni,  show.std = TRUE, pred.labels = c("Intercept", "Weekday Sleep Duration Mean (pmean)")) # OR = 2.10, CI 1.16-3.65, p=.013
```

### bootstrapping a bayesian model on Raw Data

```{r}
# reduced dataframe
merge_se_select_scaled_reduced <-
  merge_se_select_scaled %>%
  dplyr::select(
    ID, start_date_week_wednesday, timeinbed_mean_school_night_pmean, timeinbed_sd_school_night_pmean, week_num_time_z_rescaled, event_cat1
  )  %>%
  dplyr::filter(!is.na(timeinbed_mean_school_night_pmean)) %>%
  dplyr::filter(!is.na(timeinbed_sd_school_night_pmean)) %>%
  dplyr::mutate(
    event_cat1 = ifelse(event_cat1 == 1, "Event", "NonEvent"),
    event_cat1 = as.factor(event_cat1),
    event_num = ifelse(event_cat1 == "Event", 1, 0),
    ID = as.factor(ID)
    ) %>%
  dplyr::select(-event_cat1, -start_date_week_wednesday)
table(merge_se_select_scaled_reduced$event_num) # 18/2683 events
```


1) Repeats a bootstrap B (100) times. 2) Each iteration resamples participants with replacement. 3) Fits a Bayesian GLMM (brm()) to that resampled data. 4) Extracts model estimates and diagnostics. 5) Stores everything in a list.
```{r}
library(dplyr)
library(brms)
library(posterior)
library(furrr)


set.seed(123)

# Number of bootstrap replicates
B <- 100  

ids <- unique(merge_se_select_scaled_reduced$ID)

# Plan parallelization for efficiency
plan(multisession, workers = 4)  # of CPU cores

all_terms <- c("Intercept", 
               "timeinbed_mean_school_night_pmean", 
               "timeinbed_sd_school_night_pmean", 
               "week_num_time_z_rescaled")

priors <- c(
  set_prior("normal(0,1)", class = "b"),
  set_prior("student_t(3,0,2.5)", class = "sd"),
  set_prior("student_t(3,0,2.5)", class = "Intercept")
)

boot_results <- future_map(1:B, function(bi) { # future_map loops over the sequence 1:B (will perform 100 bootstrap iterations) in parallel. The function(bi) argument is just an anonymous function (no name) that will tell us which bootstrap replicate we're on. For each bi (1:B), the function resamples IDs, fits the bayesian model, extracts diagnostics, and returns them. These results are collected into a list.
  # 1) resample IDs
  boot_ids <- sample(ids, length(ids), replace = TRUE) # randomly resamples the participant IDs with replacement (some IDs appear multiple times, others not at all).
  boot_raw <- merge_se_select_scaled_reduced %>% filter(ID %in% boot_ids) # filters the full dataset (merge_se_select_scaled_reduced) to include only those resampled IDs, which forms the bootstrap dataset.

  # 2) fit bayesian glmer
  fit_b <- tryCatch( # tryCatch allows the code to continue running. So, if an iteration of brm() fails, it'll just return a NULL for that iteration (see error = function(e) NULL)
    brm(
      event_num ~ timeinbed_mean_school_night_pmean + timeinbed_sd_school_night_pmean + week_num_time_z_rescaled + (1 | ID),
      data = boot_raw,
      family = bernoulli(),
      prior = c(
        set_prior("normal(0,2)", class="b"),
        set_prior("student_t(3,0,2.5)", class="sd")
      ), # MCMC settings below (1000 iterations for each chain)
      chains = 2, iter = 1000, refresh = 0, backend = "cmdstanr", # refresh = 0 suppresses console output and cmdstanr executes faster sampling
      control = list(adapt_delta = 0.99, max_treedepth = 15)
    ), # adapt delta 0.99 reduces divergent transitions, and max_treedepth = 15 allows longer HMC trajectories
    error = function(e) NULL
  )
  
  if (!is.null(fit_b)) { # If model fit succeeds, then extract fixed effects, diagnostics, counts divergences, etc. and return all results
    
    # extract fixed effects
    est <- fixef(fit_b, summary = TRUE)[, "Estimate"] # extracts the posterior mean estimates of the fixed effects
    est_full <- setNames(rep(NA, length(all_terms)), all_terms) # ensures all predictors are returned
    est_full[names(est)] <- est # fills in the actual estimates for the terms that exist
    
    # extract diagnostics
    draws_arr <- as_draws_array(fit_b) # Converts posterior samples into an array
    
    # Rhat and ESS from posterior package
    rhat_vals <- rhat(draws_arr) # checks convergence between chains (<1.01 is good).
    ess_bulk_vals <- ess_bulk(draws_arr) # effective sample size for central and tail posterior regions. Lower ESS means there was poorer mixing
    ess_tail_vals <- ess_tail(draws_arr)
    
    # divergences (using cmdstanr)
    divs <- if (!is.null(fit_b$diagnostics$Sampler)) {
      sum(fit_b$diagnostics$Sampler[, "divergent__"]) # If diagnostics are available, it counts the number of divergent transitions; otherwise returns NA.
    } else {
      NA_integer_
    }
    
    # posterior evidence (P(β > 0))
    draws_df <- as_draws_df(fit_b) # converts the fitted model's posterior samples (fit_b) into a df where each col corresponds to a model parameter, and each row corresponds to one posterior draw (sample) from the MCMC chains. So, each column represents posterior uncertainty about a parameter.
    beta_terms <- names(draws_df)[grepl("^b_", names(draws_df))]  # fixed-effect terms only, names returns all column/parameter names
    
    ppos_vals <- sapply(beta_terms, function(term) { # loops over each beta_term fixed effect
      mean(draws_df[[term]] > 0, na.rm = TRUE) # draws_df[[term]] extracts all posterior draws for that parameter, and the mean computes the proportion of draws greater than 0, which is the posterior probability that the effect is positive P(β>0∣data).
    }) # ppos_vals is a vector with each term and corresponding posterior probability
    
    ppos_full <- setNames(rep(NA, length(all_terms)), all_terms) # outputs a vector of NA values with names matching all the model terms you expect from all_terms; if some parameter failed to estimate, the output will always include the same set of term names 

    names_map <- gsub("^b_", "", names(ppos_vals))
    ppos_full[names_map] <- ppos_vals # filling in ppos_full with the probabilities in ppos_vals
    
    return(list( # saves results into a list - one entry per bootstrap iteration
      estimates = est_full,
      rhat = rhat_vals,
      ess_bulk = ess_bulk_vals,
      ess_tail = ess_tail_vals,
      divergences = divs,
      p_pos = ppos_full # evidence probabilities      
    ))
    
  } else {
    # If model fit failed: return NAs
    return(list(
      estimates = setNames(rep(NA, length(all_terms)), all_terms),
      rhat = NA, ess_bulk = NA, ess_tail = NA, divergences = NA
    ))
  }
  
}, .progress = TRUE, .options = furrr_options(seed = TRUE))

# Extract just estimates from each replicate
estimates_list <- lapply(boot_results, `[[`, "estimates") # applies the boot_results function to each element of the list


boot_mat <- do.call(rbind, estimates_list) # converts the estimates list into a matrix (each row is a bootstrap replicate, each column is a parameter estimate)

dim(boot_mat)
head(boot_mat)

ppos_list <- lapply(boot_results, `[[`, "p_pos")
ppos_mat  <- do.call(rbind, ppos_list)

boot_mat_clean <- boot_mat[complete.cases(boot_mat), ] # returns complete cases
ppos_mat_clean  <- ppos_mat[complete.cases(ppos_mat), ]


boot_summary <- apply(boot_mat_clean, 2, function(x) { # applies a function across columns of a matrix, which is each parameter's bootstrap estimate)
  c(
    q025   = quantile(x, 0.025, na.rm = TRUE),
    median = median(x, na.rm = TRUE),
    q975   = quantile(x, 0.975, na.rm = TRUE)
  )
})

boot_summary <- t(boot_summary) # transpose
boot_summary

# Posterior evidence summary (stability of effect sign)
ppos_summary <- apply(ppos_mat_clean, 2, function(x) {
  c(
    median_p_pos = median(x, na.rm = TRUE),
    q025_p_pos   = quantile(x, 0.025, na.rm = TRUE),
    q975_p_pos   = quantile(x, 0.975, na.rm = TRUE),
    prop_strong  = mean(x > 0.95, na.rm = TRUE)
  )
})
ppos_summary <- t(ppos_summary)


# Rhat
rhat_list <- lapply(boot_results, `[[`, "rhat")
rhat_mat  <- do.call(rbind, rhat_list)

# ESS (bulk & tail)
ess_bulk_list <- lapply(boot_results, `[[`, "ess_bulk")
ess_tail_list <- lapply(boot_results, `[[`, "ess_tail")

ess_bulk_mat <- do.call(rbind, ess_bulk_list)
ess_tail_mat <- do.call(rbind, ess_tail_list)

# Divergences
divergences <- sapply(boot_results, `[[`, "divergences") # simplifies the result into a vector/matrix

# Summarize Rhat across bootstraps (per parameter)
rhat_summary <- apply(rhat_mat, 2, function(x) {
  c(
    mean   = mean(x, na.rm = TRUE),
    median = median(x, na.rm = TRUE),
    max    = max(x, na.rm = TRUE)
  )
})

# Summarize ESS (bulk and tail)
ess_bulk_summary <- apply(ess_bulk_mat, 2, median, na.rm = TRUE)
ess_tail_summary <- apply(ess_tail_mat, 2, median, na.rm = TRUE)

# Divergences (overall % of runs with any divergences)
divergence_rate <- mean(divergences > 0, na.rm = TRUE)

###  ^^^ gives you bootstrap distributions + Bayesian posteriors, so you can see both within-model uncertainty and across-sample variability.

### Now we have a boot_mat (a matrix of bootstrapped coefficient means). Can convert into a table
boot_long <- as.data.frame(boot_mat_clean) %>%
  pivot_longer(cols = everything(), names_to = "term", values_to = "estimate") %>%
  mutate(OR = exp(estimate))   # add odds ratio column

boot_summary <- boot_long %>%
  group_by(term) %>%
  summarise(
    median_logodds = median(estimate, na.rm = TRUE),
    low95_logodds  = quantile(estimate, 0.025, na.rm = TRUE),
    high95_logodds = quantile(estimate, 0.975, na.rm = TRUE),
    median_OR = median(OR, na.rm = TRUE),
    low95_OR  = quantile(OR, 0.025, na.rm = TRUE),
    high95_OR = quantile(OR, 0.975, na.rm = TRUE),
    .groups = "drop"
  )

# Compute bootstrap CIs
 boot_summary %>%
   dplyr::mutate(
      term = dplyr::recode(
        term,
        'Intercept' = 'Intercept',
        'timeinbed_mean_school_night_pmean' = 'Within-Person Average \n Sleep Duration \n  on Weekdays (Mean)',
        'timeinbed_sd_school_night_pmean' = 'Within-Person Variability \n of Sleep Duration \n on Weekdays (Mean)',
        'week_num_time_z_rescaled' = 'Mean Week Number'
        )
      ) %>%
   ggplot(
     aes(x = term, y = median_OR)
   ) +
  geom_pointrange(aes(ymin = low95_OR, ymax = high95_OR), color = "#F5B041") +
  geom_hline(yintercept = 1, linetype = "dashed", color = "#999999") +
  coord_flip() +
  scale_y_log10() +  # log scale to handle wide OR ranges
  theme_classic() +
  theme(
    strip.text.y = element_blank(),
    axis.text.x = element_text(colour = "black"),
    axis.text.y = element_text(colour = "black") 
    ) +
  labs(
    title = "Bootstrap 95% CIs for Odds Ratios",
    y = "Odds Ratio (log scale)",
    x = "Coefficient"
    ) 
ggsave("or_se_plot_bootstrapped.jpeg", width = 5, height = 3)
# OR = 1 means no effect (the dashed red line).
# If the 95% CI doesn’t cross 1, we can maybe conclude that the predictor has a robust effect across bootstraps.

# If we want to visualize posterior evidence stability
library(ggplot2)
ppos_long <- as.data.frame(ppos_mat_clean) %>%
  pivot_longer(cols = everything(), names_to = "term", values_to = "p_pos")

ppos_long %>%
   dplyr::mutate(
      term = dplyr::recode(
        term,
        'Intercept' = 'Intercept',
        'timeinbed_mean_school_night_pmean' = 'Within-Person Average \n Sleep Duration \n  on Weekdays (Mean)',
        'timeinbed_sd_school_night_pmean' = 'Within-Person Variability \n of Sleep Duration \n on Weekdays (Mean)',
        'week_num_time_z_rescaled' = 'Mean Week Number'
        )
      ) %>%
  ggplot(
    aes(x = p_pos, fill = term)
  ) +
  geom_boxplot() +
  geom_vline(xintercept = 0.95, linetype = "dashed") + # The proportion of bootstraps where evidence was “strong” (likely p_pos > 0.95).
  labs(
    x = expression(paste("Posterior probability ", beta, " > 0")),
    y = "Bootstrap count",
    title = "Stability of effect evidence across bootstraps"
  ) +
  scale_fill_manual(values = colorpalette) +
  theme_classic() + 
  theme(
    strip.text.y = element_blank(),
    axis.text.x = element_text(colour = "black"),
    axis.text.y = element_text(colour = "black") 
    )
ggsave("stability_bootstrap.jpeg", width = 5, height = 3)

ppos_summary_df <-  as_tibble(ppos_summary, rownames = "term")

ppos_summary_df <- 
  ppos_summary_df %>%  
mutate(
    term = recode(
      term,
      'Intercept' = 'Intercept',
      'timeinbed_mean_school_night_pmean' = 'Within-Person Average \nSleep Duration (Weekdays)',
      'timeinbed_sd_school_night_pmean' = 'Within-Person Variability \nSleep Duration (Weekdays)',
      'week_num_time_z_rescaled' = 'Mean Week Number'
    )
  ) %>%
  ggplot(aes(
    x = term,
    y = median_p_pos
  )) +
  geom_pointrange(
    aes(
      ymin = `q025_p_pos.2.5%`,
      ymax = `q975_p_pos.97.5%`
    ),
    color = "#2E86AB",
    size = 1
  ) +
  geom_hline(yintercept = 0.95, linetype = "dashed", color = "#999999") +
  geom_hline(yintercept = 0.5, linetype = "dotted", color = "#BBBBBB") +
  coord_flip() +
  scale_y_continuous(limits = c(0, 1), labels = scales::percent) +
  theme_classic(base_size = 14) +
  labs(
    title = "Posterior Probability of Positive Effect (p_pos)",
    subtitle = "Across bootstrapped Bayesian GLMMs",
    x = "",
    y = "Probability(β > 0)",
    caption = "Dashed line = 0.95 threshold for 'strong evidence'"
  )
ppos_summary_df
```
