---
title: "3_sleep_and_si_lagged_weekly_associations_cont"
author: "Jackie Kirshenbaum"
date: "`r format(Sys.Date(), '%B %e %Y')`"
output:
  html_document:
    df_print: paged
  toc: yes
  toc_float: yes
---

This script examines lagged week-to-week associations between SI (continuous) and sleep features (time-in-bed/dur, bedtime)

# Libraries and Themes

```{r}
library(tidyverse)
library(scipub)
library(kableExtra)
library(skimr)
library(lmerTest)
library(brms)
library(naniar)
library(sjPlot)
library(broom)
library(ggeffects)
library(patchwork)
library(jtools)
library(performance)
library(interactions)
library(ggplot2)

theme_set(theme_classic()); theme_update(axis.title.x = element_text(face = "bold"),axis.title.y = element_text(face = "bold"))

colorpalette <- c("#F17E0B", "#F5B7B1", "#999999","#BABDE1", "#F5B041", "#B5EAF0", "#CBD1D4",  "#CFE4EF") 
```

# Load data

```{r}
# weekly with suicide risk data
load("/../../../../Jackie/sleep_and_suicide_2024-2025/weekly_sleep_and_stb_and_sr_cleaned_March272026.Rda")
```

# Scaling variables

```{r}
#  person centering and group centering variables
merge_si_select_lag_scaled <- 
  merge_si_select_lag %>%
  dplyr::ungroup() %>%
  dplyr::group_by(ID) %>% # p-mean centering (creating within-person / idiographic variables)
  dplyr::mutate(
    # tib weekly
    timeinbed_whole_week_mean_pmean = as.numeric(scale(timeinbed_whole_week_mean)), # person's mean duration that week relative to their mean duration on other weeks 
    timeinbed_mean_school_night_pmean = as.numeric(scale(timeinbed_mean_school_night)),
    timeinbed_mean_free_night_pmean = as.numeric(scale(timeinbed_mean_free_night)),
    timeinbed_whole_week_sd_pmean = as.numeric(scale(timeinbed_whole_week_sd)),   # person's variability in duration that week relative to their mean variability in duration on other weeks    
    timeinbed_sd_school_night_pmean = as.numeric(scale(timeinbed_sd_school_night)),      
    timeinbed_sd_free_night_pmean = as.numeric(scale(timeinbed_sd_free_night)),
    # bedtime weekly
    bedtime_whole_week_mean_pmean = as.numeric(scale(bedtime_whole_week_mean)),  
    bedtime_mean_school_night_pmean = as.numeric(scale(bedtime_mean_school_night)),
    bedtime_mean_free_night_pmean = as.numeric(scale(bedtime_mean_free_night)),
    bedtime_mean_free_night_pmean_hrs = as.numeric(scale(bedtime_mean_free_night/60, scale = FALSE)),  
    bedtime_whole_week_sd_pmean = as.numeric(scale(bedtime_whole_week_sd)),
    bedtime_sd_school_night_pmean = as.numeric(scale(bedtime_sd_school_night)),
    bedtime_sd_free_night_pmean = as.numeric(scale(bedtime_sd_free_night)),
    bedtime_sd_free_night_pmean_hrs = as.numeric(scale(bedtime_sd_free_night/60, scale = FALSE)),  
    # weekly mood
    mood_whole_week_mean_pmean = as.numeric(scale(mood_whole_week_mean)),
    mood_R_whole_week_mean_pmean = as.numeric(scale(mood_R_whole_week_mean)),
    mood_mean_free_night_pmean = as.numeric(scale(mood_mean_free_night)),
    mood_R_mean_free_night_pmean = as.numeric(scale(mood_R_mean_free_night)),
    mood_mean_school_night_pmean = as.numeric(scale(mood_mean_school_night)),
    mood_R_mean_school_night_pmean = as.numeric(scale(mood_R_mean_school_night)),
  ) %>%
  dplyr::ungroup() %>% 
  dplyr::mutate(
    # sex
    sex = factor(sex),
    # age
    age_z = as.numeric(scale(age)),
    week_num_time_z_rescaled = as.numeric(scale(week_num)),
    schoolclosure = factor(schoolclosure),
    summer = factor(summer),
    # SR variables
    SSI5_i_z = as.numeric(scale(SSI5_i)),
    MFQ_i_z = as.numeric(scale(MFQ_i)),
    SCARED_i_z = as.numeric(scale(SCARED_i)),
    SHAPS_i_z = as.numeric(scale(SHAPS_TOT_r_i)),
    MEQ_total_i_z = as.numeric(scale(MEQ_total_i)),
    psqitotal_z = as.numeric(scale(psqitotal)),
    weekly_missingness_z = as.numeric(scale(missing_lead_weeks)),    
    risk_SIsig4_future = as.factor(risk_SIsig4_future),
    risk_SIsig3_future = as.factor(risk_SIsig3_future),
    risk_SInum_future_re = case_when(
    risk_SInum_future == 1 ~ 0,
    risk_SInum_future == 2 ~ 1,
    risk_SInum_future == 3 ~ 2,
    risk_SInum_future == 4 ~ 3,
    risk_SInum_future == 5 ~ 4,    
    TRUE ~ risk_SInum_future # Default condition for all other values
    )
  ) %>%
  # dplyr::select(ID:event_bin, GROUP, everything())  %>%
  dplyr::mutate(across(ends_with("pmean"), ~ifelse(is.nan(.), NA, .)))
# weekly_sleep_and_stb_and_sr_cleaned_March272026.Rda
```

__Q2:__ Which sleep parameters predict week-to-week suicidal ideation? (continuous outcome)

```{r}
options(scipen = 999)
```

# Distribution of SI
```{r}
merge_si_select_lag_scaled %>%
  ggplot(
    aes(x = risk_SInum_future)
  ) +
  geom_histogram() +
#  geom_freqpoly(bins = 30) +
  theme(
    geom = element_geom(
    bordertype = "dashed",
    borderwidth = 0.2,
    linewidth = 2,
    linetype = "solid"
    )
  ) +
  labs(
    x = "Frequency of Suicidal Ideation Severity \n 1=never; 5=all of the time"
  )
ggsave("FigureX_SI_distribution.jpeg", width = 5.5, height = 4)
```

# Is a ZIP needed for modeling SI as a continuous outcome?

```{r}
mean(merge_si_select_lag_scaled$risk_SInum_future_re == 0, na.rm = TRUE) # Yes, > 50% 0s
```

# Covariates?
Testing potential covariates in relation to my DV

```{r}
# individual
cov_age_sex <- lmer(risk_SInum_future ~ age_z + sex + (1 | ID),  data = merge_si_select_lag_scaled)
summary(cov_age_sex) # age and sex both significant

cov_sch_close <- lmer(risk_SInum_future ~ schoolclosure + (1 | ID), data = merge_si_select_lag_scaled) 
summary(cov_sch_close) # significant

cov_summ <- lmer(risk_SInum_future ~ summer + (1 | ID), data = merge_si_select_lag_scaled)
summary(cov_summ)

cov_week <- lmer(risk_SInum_future ~ week_num_time_z_rescaled + (1 | ID), data = merge_si_select_lag_scaled) 
summary(cov_week)

cov_phone <- lmer(risk_SInum_future ~ iPhone +  (1 | ID), data = merge_si_select_lag_scaled)
summary(cov_phone)

cov_ssi <- lmer(risk_SInum_future ~  SSI5_i_z + (1 | ID), data = merge_si_select_lag_scaled) # ssi sig
summary(cov_ssi)

cov_mfq <- lmer(risk_SInum_future ~ MFQ_i_z +  (1 | ID), data = merge_si_select_lag_scaled) # mfq sig
summary(cov_mfq)

cov_meq <- lmer(risk_SInum_future ~ MEQ_total_i_z +   (1 | ID), data = merge_si_select_lag_scaled) 
summary(cov_meq)

cov_scared <- lmer(risk_SInum_future ~ SCARED_i_z + (1 | ID), data = merge_si_select_lag_scaled) 
summary(cov_scared)

cov_shaps <- lmer(risk_SInum_future ~ SHAPS_i_z + (1 | ID), data = merge_si_select_lag_scaled) # shaps sig
summary(cov_shaps)

cov_psqi <- lmer(risk_SInum_future ~ psqitotal_z + (1 | ID), data = merge_si_select_lag_scaled) 
summary(cov_psqi)

cov_mood <- lmer(risk_SInum_future ~ mood_whole_week_mean_pmean + (1 | ID), data = merge_si_select_lag_scaled)
summary(cov_mood)

cov_mood_school <- lmer(risk_SInum_future ~ mood_mean_school_night_pmean + (1 | ID), data = merge_si_select_lag_scaled)
summary(cov_mood_school)

cov_mood_free <- lmer(risk_SInum_future ~ mood_mean_free_night_pmean + (1 | ID), data = merge_si_select_lag_scaled)
summary(cov_mood_free)

tab_model(cov_age_sex, cov_sch_close, cov_summ, cov_week, cov_phone, cov_ssi, cov_mfq, cov_meq, cov_scared, cov_shaps, cov_psqi, cov_mood, cov_mood_school, cov_mood_free)

# unique variance?
cov <- lmer(risk_SInum_future ~ age_z + sex + schoolclosure + SSI5_i_z + MFQ_i_z + SHAPS_i_z + (1 | ID), data = merge_si_select_lag_scaled) 
summary(cov)

tab_model(cov, pred.labels = c("Intercept", "Age", "Sex", "School Closure", "Baseline SI", "Baseline Depression", "Baseline Anhedonia"))
# only SSI is uniquely associated with clinically sig SI
```

##### Supplementary Analysis of Missingness #####

```{r}
missingness_m1 <- lmer(risk_SInum_future ~ weekly_missingness_z + (1 | ID), data = merge_si_select_lag_scaled)
summary(missingness_m1)
tab_model(missingness_m1, show.std = TRUE)
```


# Person-Centered Logistic Reg Whole Week Mean + SD

```{r}
#### TIB (whole week)

log_tib_step1 <- lmer(risk_SInum_future_re ~ timeinbed_whole_week_sd_pmean + timeinbed_whole_week_mean_pmean  + SSI5_i_z + schoolclosure + (1 | ID), data = merge_si_select_lag_scaled)
summary(log_tib_step1)


tab_model(log_tib_step1, show.std = TRUE, pred.labels = c("Intercept", "Weekly TIB SD (pmean)", "Weekly TIB Mean (pmean)", "Baseline SI", "School Closure"))

#### BEDTIME (whole week)

log_on_step1 <- lmer(risk_SInum_future ~ bedtime_whole_week_sd_pmean + bedtime_whole_week_mean_pmean + SSI5_i_z + schoolclosure + (1 | ID), data = merge_si_select_lag_scaled)
summary(log_on_step1)

tab_model(log_on_step1, show.std = TRUE, pred.labels = c("Intercept", "Weekly Bedtime SD (pmean)", "Weekly Bedtime Mean (pmean)", "Baseline SI", "School Closure"))
```


###################### WEEKEND / WEEKDAY ############################

## TIB Weekdays

```{r}
log_tib_step1_ww <- lmer(risk_SInum_future ~ timeinbed_mean_school_night_pmean + timeinbed_sd_school_night_pmean + SSI5_i_z  + schoolclosure + (1 | ID), data = merge_si_select_lag_scaled)
summary(log_tib_step1_ww) 

tab_model(log_tib_step1_ww,  show.std = TRUE, pred.labels = c("Intercept", "Weekday TIB Mean (pmean)", "Weekday TIB SD (pmean)", "Baseline SI", "School Closure"))

```

## TIB Weekends

```{r}
log_tib_step1_wwend <- lmer(risk_SInum_future ~ timeinbed_mean_free_night_pmean + timeinbed_sd_free_night_pmean + SSI5_i_z + schoolclosure + (1 | ID), data = merge_si_select_lag_scaled)
summary(log_tib_step1_wwend) 

tab_model(log_tib_step1_wwend, show.std = TRUE, pred.labels = c("Intercept", "Weekend TIB Mean (pmean)", "Weekend TIB SD (pmean)", "Baseline SI", "School Closure"))

```

## Bedtime Weekdays

```{r}
log_on_step1_ww <- lmer(risk_SInum_future ~ bedtime_mean_school_night_pmean + bedtime_sd_school_night_pmean + SSI5_i_z  + schoolclosure + (1 | ID), data = merge_si_select_lag_scaled)
summary(log_on_step1_ww) 

tab_model(log_on_step1_ww, show.std = TRUE, pred.labels = c("Intercept", "Weekday Bedtime Mean (pmean)", "Weekday Bedtime SD (pmean)", "Baseline SI", "School Closure"))
```

## Bedtime Weekends

```{r}
log_on_step1_wwend <- lmer(risk_SInum_future ~ bedtime_mean_free_night_pmean + bedtime_sd_free_night_pmean + SSI5_i_z + schoolclosure + (1 | ID), data = merge_si_select_lag_scaled)
summary(log_on_step1_wwend) 

tab_model(log_on_step1_wwend, show.std = TRUE, pred.labels = c("Intercept", "Weekend Bedtime Mean (pmean)", "Weekend Bedtime SD (pmean)", "Baseline SI", "School Closure"))

```

# Overall Models

```{r}
tab_model(log_on_step1, log_on_step1_ww, log_on_step1_wwend, show.std = TRUE, pred.labels = c("Intercept", "Whole-Week Bedtime SD (pmean)", "Whole-Week Bedtime Mean (pmean)", "Baseline Suicidal Ideation Severity", "School Closure", "Weekday Bedtime Mean (pmean)", "Weekday Bedtime SD (pmean)", "Weekend Bedtime Mean (pmean)", "Weekend Bedtime SD (pmean)"))

tab_model(log_tib_step1, log_tib_step1_ww, log_tib_step1_wwend, show.std = TRUE, pred.labels = c("Intercept", "Whole-Week Sleep Duration SD (pmean)", "Whole-Week TIB Mean (pmean)", "Baseline Suicidal Ideation Severity", "School Closure", "Weekday TIB Mean (pmean)", "Weekday TIB SD (pmean)", "Weekend TIB Mean (pmean)", "Weekend TIB SD (pmean)"))
```

Overall, results indicate that a greater bedtime inconsistency relative to one's typical bedtime variability on a given weekend is related to next-week SI.

